### 2020-07-21 Cluster Analysis.Rmd
commit 2

---
title: "Cluster Analysis"
author: "Nadia Noui-Mehidi"
date: "7/15/2020"
output: html_document
---

```{r setup, include=FALSE}
college <- read.csv(file="file:///Users/nadianoui-mehidi/Desktop/college_acceptance.csv",head=TRUE,sep=",")

recidivism <- read.csv(file="file:///Users/nadianoui-mehidi/Desktop/recidivism.csv",head=TRUE,sep=",")
library(dplyr)
library(skimr)
library(cluster)
require(tidyverse)
library(kohonen)
require(ggplot2)
require(ggridges)
require(RColorBrewer)
library(ggplot2)
library(vegan)

```

#COMPONENT 1:    Multidimensional Scaling
##1. EDA
The recidivism dataset is an 18 variable dataset with 1445 records.  The data is a random sample of convicts released from prison between July 1, 1977 and June 30, 1978. The information was collected retrospectively by looking at records in April 1984, so the maximum possible length of observation is 81 months.

The 18 variables including in our data set are:
 1. black =1 if black
 2. alcohol =1 if alcohol problems
 3. drugs =1 if drug history
 4. supervised =1 if release supervised
 5. married =1 if married when incarc.
 6. felony =1 if felony sentence
 7. workprg =1 if in N.C. pris. work prg.
 8. property =1 if property crime
 9. person =1 if crime against person
10. nbr_piors # prior convictions
11. education years of schooling
12. nbr_rules # rules violations in prison
13. age in months
14. time_served time served, rounded to months
15. follow_up length follow period, months
16. duration max(time until return, follow)
17. censored =1 if duration right censored
18. log_duration log(duration)

```{r}
glimpse(recidivism)
```
There are 9 binary, 8 interval and 1 continuous variable in our dataset. We normally dont want to include categorical variables when multidemenisional scaling but since our categorical variables are binary their distance measures are still meaningful and we can keep them in.

Of interest to us is the time until they return to prison (duration variable), so with this in mind, we will keep the following variables:  
duration:months until return to prison
censored:censoring indicator variable
workprg: indicator of participation in a work program
nbr_priors: the number of previous convictions 
time_served: the time served rounded to months 
felony: an indicator of felony sentences
alcohol: an indicator of alcohol problems
drugs: an indicator of drug use history
black: an indicator for African Americans 
married: an indicator if married when incarcerated 
education: the number of years of schooling 
age: in months.

We created a new dataframe with only the variables we are using. Since our variables have different scales, we also standardize the records so that all variables have an equal opportunity to influence the patterns.

```{r}
df = subset(recidivism, select = -c(super, property, person, rules, follow, ldurat))
```

## Including Plots

You can also embed plots, for example:

```{r pressure, echo=FALSE}
plot(pressure)
```

Note that the `echo = FALSE` parameter was added to the code chunk to prevent printing of the R code that generated the plot.